Students' Alcohol Consumption impacts on academic Results on the example of a secondary Scool Maths Class in Portugal

### 2016-10-15 Working-Files/WF_MSprojectReport before knitting tests.Rmd
Add files via upload

# Students' Alcohol Consumption impacts on academic Results on the example of a secondary Scool Maths Class in Portugal

@@

2. Dataset Description

3. Descriptive Statistics

4. Correlation Analysis

5. Hypothesis Testing

6. Regression

7. Conclusion

8. References

9. Appendices
---

## Introduction

Recent events in the world economy have lead to the decline in economic performance in many countries, amongst which is the decline of econommic growth in the European Union (EU).  As a response to these changes,  the EU has made it a priority to reverse this trend through investments that would provide new jobs and create a better educated workforce. Hence, the European Commission (EC) has set up a ‘Strategy for smart, sustainable and inclusive growth’ (European Comission, 2010), where a crucial part of investment is to contribute to the education of the next generations: The aim is to reduce the share of early school leavers to below 10% and ensure that at least 40% of the younger generation obtain a tertiary degree.

@@

## Description of Dataset and Descriptive Statistics

```{r pressure, include=FALSE}
sm <- read.csv("student-mat.csv", sep=';')

# Descriptive statistics
library(ggplot2)

@@

df_gender <- data.frame(Gender=c("Female", "Male"), value=c(208, 187))
pie_gender <- ggplot(data=df_sex, aes(x="", y=value, fill=Gender)) + 
  geom_bar(width = 1, stat = "identity") + 
  coord_polar("y", start=0) + 
  scale_fill_brewer() + blank_theme + 
  theme(axis.text.x=element_blank()) +
  geom_text(aes(y = value/3 + c(0, cumsum(value)[-length(value)]), label = (value)), size=4)

@@

df_studyTime <- data.frame(Study_Time=c('<2 hours', "2 to 5 hours", "5 to 10 hours", ">10 hours"), value=c(105, 198, 65, 27))
#weekly study time (numeric: 1 - <2 hours, 2 - 2 to 5 hours, 3 - 5 to 10 hours, or 4 - >10 hours)

bar_studytime <- ggplot(data=df_studyTime, aes(x="", y=value, fill=Study_Time)) + 
  geom_bar(width = 1, stat = "identity") + 
  scale_fill_brewer() + blank_theme + 
  theme(axis.text.x=element_blank()) +
  geom_text(aes(y = value/3 + c(0, cumsum(value)[-length(value)]), label = (value)), size=4)

@@

df_school <- data.frame(Schools=c("Gabriel Pereira", "Mousinho da Silveira"), value=c(349,46))
bar_school <- ggplot(data=df_school, aes(x="", y=value, fill=Schools)) + 
  geom_bar(width = 1, stat = "identity") + 
  blank_theme + 
  theme(axis.text.x=element_blank()) +
  geom_text(aes(y = value/3 + c(0, cumsum(value)[-length(value)]), label = (value)), size=4)

@@

df_fjob <- data.frame(Group = c("Health", "Work at Home", "Teacher", "Servies", "Other"), value = c(18, 20, 29, 111, 217))
pie_fjob <- ggplot(data = df_fjob, aes(x="", y=value, fill=Group)) + 
  geom_bar(width = 1, stat = "identity") + 
  coord_polar("y", start=0) + 
  scale_fill_brewer() + blank_theme +
  theme(axis.text.x=element_blank()) +
  geom_text(aes(y = value/3 + c(0, cumsum(value)[-length(value)]), label = (value)), size=4) +
  ggtitle("Father's Job")

@@

df_mjob <- data.frame(Group = c("Health", "Teacher", "Work at Home", "Services", "Other"), value = c(34, 58, 59, 103, 141))
pie_mjob <- ggplot(data = df_mjob, aes(x="", y=value, fill=Group)) + 
  geom_bar(width = 1, stat = "identity") + 
  coord_polar("y", start=0) + 
  scale_fill_brewer() + blank_theme +
  theme(axis.text.x=element_blank()) +
  geom_text(aes(y = value/3 + c(0, cumsum(value)[-length(value)]), label = (value)), size=4) +
  ggtitle("Mother's Job")

@@

df_guardian <- data.frame(group = c("Father", "Mother", "Other"), value = c(90, 273, 32))
pie_guardian <- ggplot(data = df_guardian, aes(x="", y=value, fill=group)) + 
  geom_bar(width = 1, stat = "identity") + 
  coord_polar("y", start=0) + 
  scale_fill_brewer() + blank_theme +
  theme(axis.text.x=element_blank()) +
  geom_text(aes(y = value/3 + c(0, cumsum(value)[-length(value)]), label = (value)), size=4) +
  ggtitle("Guardian")

###############################################################################################

@@

#### 1. General Information
General information about the students include gender, address and age. From the descriptive statistics (section 3) we can derive that the age from the students in this dataset ranges from 15 to 22 years with an average of 16.7 years. Out of the total of 395, there are 52.7% female and 47.3% male students. Most students would live in an urban environment, while approx. 20% live in an rural environment.

```{r, echo=FALSE}
pie_gender

pie_address
```

#### 2. Family  
Information in the students families includes the family size, educational background of their parents and their jobs, and the student's guardian.

```{r, echo=FALSE}
pie_fjob

pie_mjob

pie_guardian
```

#### 3. Education
The education section of this dataset covers the split between the two schools observed, why each student has chosen that particular school, and how long it takes each student to get to school. In addition, it provides information on the student's study time, extra educational support or classes taken, whether they have failed to pass any classes and if the student is involved in extra curricular activities. Finally, there is data on their previous education and aspirations on whether to go into higher education.

```{r, echo=FALSE}
bar_school
bar_studytime

#### 4. Social Life and Health 
The social life section provieds information on activities the student is involved in outside of school. This includes

#### 5. Grades
The grades are split into three grades: the first period (G1), second period (G2) and final grade (G3). For the purpose of thy analysis, G1 and G2 will be neglected. 
```{r, echo=FALSE}
pie_failures

## Correlation Analysis

Text

```{r pressure, echo=FALSE}
plot(pressure)
```
```{r}
sm <- student.mat
for (i in 16:24) {                                  
  sm[,i] <- as.character(sm[,i])
  for (j in 1:nrow(sm)) {
    if (sm[j,i] == "yes") sm[j,i] <- 1
    if (sm[j,i] == "no") sm[j,i] <- 0
  }
  sm[,i] <- as.numeric(sm[,i])
}

#General test 
```{r}
t.test(sm[,"Walc"], sm[,"Dalc"])
library(ggplot2)
ggplot(data = sm) + geom_bar(aes(x=age))
ggplot(data = sm) + geom_bar(aes(x=sex))
ggplot(data = sm) + geom_bar(aes(x=Dalc))
ggplot(data = sm) + geom_bar(aes(x=Walc))
i <- ggplot(sm, aes(x = sm[,"age"], y = sm[,"sex"]))
i + geom_bin2d(binwidth = c(1,1)) #LABEL #Distribution of gender and age
i <- ggplot(sm, aes(x = sm[,"Dalc"], y = sm[,"Walc"]))
i + geom_bin2d(binwidth = c(1,1)) #LABEL #Distribution of Walc and Dalc

#Test and plot on gender and age

###age
```{r}
f <- ggplot(data =sm, aes(x = sm[,"Dalc"], y = sm[,"age"]))
f + geom_smooth(method = lm) #LABEL
f <- ggplot(data =sm, aes(x = sm[,"Walc"], y = sm[,"age"]))
f + geom_density(method = lm)  #LABEL
f + geom_smooth(na.rm = TRUE)
i <- ggplot(sm, aes(x = sm[,"age"], y = sm[,"sex"]))
i + geom_bin2d(binwidth = c(1,1))
i <- ggplot(sm, aes(x = sm[,"absences"], y = sm[,"Walc"]))
i + geom_bin2d(binwidth = c(1,1))
cor.test(sm[,"age"],sm[,"Dalc"])
cor.test(sm[,"age"],sm[,"Walc"])
```

###gender
```{r}
for (i in 2) {                                  
  sm[,i] <- as.character(sm[,i])
  for (j in 1:nrow(sm)) {
    if (sm[j,i] == "M") sm[j,i] <- 1
    if (sm[j,i] == "F") sm[j,i] <- 0
  }
  sm[,i] <- as.numeric(sm[,i])
}
cor.test(sm[,"sex"],sm[,"Walc"])
cor.test(sm[,"sex"],sm[,"Dalc"])
```

##Correlations tests 
As inputs have to be numeric it requires less lines to split the code that way instead of doing block per parameters analysed (Failures, paid, Medu, ... )
We could have done for loops here to calculate several correlations but it the results wasn't user friendly and doint it that way gave some useless correlations.
```{r}
sm <- as.numeric(sm)
sm1 <- subset(sm, sex == 1) #Create subsets with boys or girls
sm1 <- subset(sm, sex == 0)
```

###Romantic
```{r}
cor.test(sm[,"romantic"], sm[,"Dalc"])
cor.test(sm[,"romantic"], sm[,"Walc"])
```

###absenses
```{r}
cor.test(sm[,"absences"], sm[,"Dalc"])
cor.test(sm[,"absences"], sm[,"Walc"])
```

###Failures
```{r}
cor.test(sm[,"failures"], sm[,"Dalc"])
cor.test(sm[,"failures"], sm[,"Walc"])
```

###Boypart
```{r}
cor.test(sm1[,"failures"], sm1[,"Dalc"])
cor.test(sm1[,"failures"], sm1[,"Walc"])
```

###Medu
```{r}
cor.test(sm[,"Medu"], sm[,"Dalc"])
cor.test(sm[,"Medu"], sm[,"Walc"])
```

###Fedu
```{r}
cor.test(sm[,"Fedu"], sm[,"Dalc"])
cor.test(sm[,"Fedu"], sm[,"Walc"])
```

###Travel Time
```{r}
cor.test(sm[,"traveltime"], sm[,"Dalc"])
cor.test(sm[,"traveltime"], sm[,"Walc"])
```

###famrel
```{r}
cor.test(sm[,"famrel"], sm[,"Dalc"])
cor.test(sm[,"famrel"], sm[,"Walc"])
```

###goout
```{r}
cor.test(sm[,"goout"], sm[,"Dalc"])
cor.test(sm[,"goout"], sm[,"Walc"])
```

###internet
```{r}
cor.test(sm[,"internet"], sm[,"Dalc"])
cor.test(sm[,"internet"], sm[,"Walc"])
```

###famrel
```{r}
cor.test(sm[,"famrel"], sm[,"Dalc"])
cor.test(sm[,"famrel"], sm[,"Walc"])
```

###Grades
```{r}
cor.test(sm[,"G1"], sm[,"Dalc"])
cor.test(sm[,"G1"], sm[,"Walc"])
cor.test(sm[,"G2"], sm[,"Dalc"])
cor.test(sm[,"G2"], sm[,"Walc"])
cor.test(sm[,"G3"], sm[,"Dalc"])
cor.test(sm[,"G3"], sm[,"Walc"])
cor.test(sm[,"G3"], sm[,"goout"])
cor.test(sm[,"G3"], sm[,"failures"])
cor.test(sm[,"G3"], sm[,"absences"])
```

#EARLY SIGNS OF ACADEMIC FAILURES
###abscences
```{r}
cor.test(sm[,"absences"], sm[,"Dalc"])
cor.test(sm[,"absences"], sm[,"Walc"])
cor.test(sm1[,"absences"], sm1[,"Dalc"])
cor.test(sm1[,"absences"], sm1[,"Walc"])
cor.test(sm2[,"absences"], sm2[,"Dalc"])
cor.test(sm2[,"absences"], sm2[,"Walc"])
```

###Failures
```{r}
cor.test(sm[,"failures"], sm[,"Dalc"])
cor.test(sm[,"failures"], sm[,"Walc"])
cor.test(sm1[,"failures"], sm1[,"Dalc"])
cor.test(sm1[,"failures"], sm1[,"Walc"])
cor.test(sm2[,"failures"], sm2[,"Dalc"])
cor.test(sm2[,"failures"], sm2[,"Walc"])
```

###goout
```{r}
cor.test(sm[,"goout"], sm[,"Dalc"])
cor.test(sm[,"goout"], sm[,"Walc"])
cor.test(sm1[,"goout"], sm1[,"Dalc"])
cor.test(sm1[,"goout"], sm1[,"Walc"])
cor.test(sm2[,"goout"], sm2[,"Dalc"])
cor.test(sm2[,"goout"], sm2[,"Walc"])
```

###remi text
```{r}
layout(matrix(1:14, ncol = 7))
cdplot(sm[,"guardian"] ~ sm[,"Dalc"], data = sm, xlab = "Alcohol consumption during working days ", ylab = "Guardian", border = 1)
cdplot(sm[,"guardian"] ~ sm[,"Walc"], data = sm, xlab = "Alcohol consumption during Week-end ", ylab = "Guardian", border = 1)
cdplot(sm[,"Medu"] ~ sm[,"Dalc"], data = sm, xlab = "Alcohol consumption during working days ", ylab = "Father's education", border = 1)
cdplot(sm[,"Medu"] ~ sm[,"Walc"], data = sm, xlab = "Alcohol consumption during Week-end ", ylab = "Father's education", border = 1)
cdplot(sm[,"Fedu"] ~ sm[,"Dalc"], data = sm, xlab = "Alcohol consumption during working days ", ylab = "Mother's education", border = 1)
cdplot(sm[,"Fedu"] ~ sm[,"Walc"], data = sm, xlab = "Alcohol consumption during working Week-end ", ylab = "Mother's education", border = 1)
cdplot(sm[,"Pstatus"] ~ sm[,"Dalc"], data = sm, xlab = "Alcohol consumption during working days ")
cdplot(sm[,"Pstatus"] ~ sm[,"Walc"], data = sm, xlab = "Alcohol consumption during Week-end ")
cdplot(sm[,"famsize"] ~ sm[,"Dalc"], data = sm, xlab = "Alcohol consumption during working days ")
cdplot(sm[,"famsize"] ~ sm[,"Walc"], data = sm, xlab = "Alcohol consumption during working Week-end ")
cdplot(sm[,"Mjob"] ~ sm[,"Dalc"], data = sm, xlab = "Alcohol consumption during working days")
cdplot(sm[,"Mjob"] ~ sm[,"Walc"], data = sm, xlab = "Alcohol consumption during working Week-end")
cdplot(sm[,"famrel"] ~ sm[,"Dalc"], data = sm, xlab = "Alcohol consumption during working days")
cdplot(sm[,"famrel"] ~ sm[,"Walc"], data = sm, xlab = "Alcohol consumption during working Week-end ")
```

###remi text
```{r}
layout(matrix(1:10, ncol = 5))
cdplot(sm[,"romantic"] ~ sm[,"Dalc"], data = sm)
cdplot(sm[,"romantic"] ~ sm[,"Walc"], data = sm)
cdplot(sm[,"goout"] ~ sm[,"Dalc"], data = sm)
cdplot(sm[,"goout"] ~ sm[,"Walc"], data = sm)
cdplot(sm[,"internet"] ~ sm[,"Dalc"], data = sm)
cdplot(sm[,"internet"] ~ sm[,"Walc"], data = sm)
cdplot(sm[,"activities"] ~ sm[,"Dalc"], data = sm)
cdplot(sm[,"activities"] ~ sm[,"Walc"], data = sm)
cdplot(sm[,"freetime"] ~ sm[,"Dalc"], data = sm)
cdplot(sm[,"freetime"] ~ sm[,"Walc"], data = sm)
```

###remi text
```{r}
layout(matrix(1:8, ncol = 4))
cdplot(sm[,"failures"] ~ sm[,"Dalc"], data = sm)
cdplot(sm[,"failures"] ~ sm[,"Walc"], data = sm)
cdplot(sm[,"schoolsup"] ~ sm[,"Dalc"], data = sm)
cdplot(sm[,"schoolsup"] ~ sm[,"Walc"], data = sm)
cdplot(sm[,"studytime"] ~ sm[,"Dalc"], data = sm)
cdplot(sm[,"studytime"] ~ sm[,"Walc"], data = sm)
cdplot(sm[,"G3"] ~ sm[,"Dalc"], data = sm)
cdplot(sm[,"G3"] ~ sm[,"Walc"], data = sm)

#T
###GoOUT
```{r}
sm[,"goout"] <- as.numeric(sm[,"goout"])
sm[,"Dalc"] <- as.numeric(sm[,"Dalc"])
sm[,"Walc"] <- as.numeric(sm[,"Walc"])
cor.test(sm[,"goout"], sm[,"Dalc"])
cor.test(sm[,"goout"], sm[,"Walc"]) #Stong correlation
RegGoout <- lm (goout ~ Walc, data = sm)
RegGoout
libraryr(ggplot2)
f <- ggplot(data =sm, aes(x = sm[,"goout"], y = sm[,"Walc"]))
f + geom_smooth(method = lm) #TBS
sm[,"goout"] <- as.factor(sm[,"goout"])
sm[,"Dalc"] <- as.factor(sm[,"Dalc"])
sm[,"Walc"] <- as.factor(sm[,"Walc"])
layout(matrix(1:2, ncol = 2))
cdplot(sm[,"goout"] ~ sm[,"Dalc"], data = sm)
cdplot(sm[,"goout"] ~ sm[,"Walc"], data = sm)

###Internet
```{r}
sm[,"internet"] <- as.numeric(sm[,"internet"])
sm[,"Dalc"] <- as.numeric(sm[,"Dalc"])
sm[,"Walc"] <- as.numeric(sm[,"Walc"])
cor.test(sm[,"internet"], sm[,"Dalc"])
cor.test(sm[,"internet"], sm[,"Walc"])
```

###Activities
```{r}
sm[,"activities"] <- as.numeric(sm[,"activities"])
sm[,"Dalc"] <- as.numeric(sm[,"Dalc"])
sm[,"Walc"] <- as.numeric(sm[,"Walc"])
cor.test(sm[,"activities"], sm[,"Dalc"])
cor.test(sm[,"activities"], sm[,"Walc"])

###Freetime
```{r}
sm[,"freetime"] <- as.numeric(sm[,"freetime"])
sm[,"Dalc"] <- as.numeric(sm[,"Dalc"])
sm[,"Walc"] <- as.numeric(sm[,"Walc"])
cor.test(sm[,"freetime"], sm[,"Dalc"])
cor.test(sm[,"freetime"], sm[,"Walc"])
```

##Mean tests
```{r}
mean(sm[sm[, "Fedu"] == "5", "Dalc"]) - mean(sm[sm[, "Medu"] == "5", "Dalc"])
mean(sm[sm[, "guardian"] == "mother", "Dalc"]) - mean(sm[sm[, "guardian"] == "father", "Dalc"])

Text

```{r pressure, echo=FALSE}
plot(pressure)
```

## Correlation and Regression

In this report, we would like to find factors that are related to students academic performance. We will perform a multivariate regression analysis and use step-wise variable selection to find significant variables related to our response variable (academic performance - "Score"). Prior to the regression analysis, we want to get a general idea about the relationship between the variables. There are several things we want to inspect :

1) Distribution of response variable
if there is skewness in the distribution, we should consider transforming the variable (i.e logarithmic)

2) Correlation between response variable and each explanatory variables

3) Correlation among explanatory variables
if there is a high correlation among explanatory variables, then there is possible multicollinearity problem

@@

I. Distribution of Response Variable
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
ggplot(sm) + geom_density(aes(x=Score)) + labs(x = "Academic Score", y="% Frequency", title = "Academic Score of Secondary School Student")
```
From the graph, we can see that the distribution is slightly left-skewed. Generally, we would transform the variable to become more normally distributed. However, for the purpose of this project, we will not do any transfomation on the response variable and assume that the distribution is normal.

II. Correlation between response variable and each explanatory variables
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...)
{
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  r <- abs(cor(x, y))
  txt <- format(c(r, 0.123456789), digits = digits)[1]
  txt <- paste0(prefix, txt)
  if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
  text(0.5, 0.5, txt, cex = cex.cor * r)
}

@@

III. Correlation between response variable and each explanatory variables

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...)
{
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  r <- abs(cor(x, y))
  txt <- format(c(r, 0.123456789), digits = digits)[1]
  txt <- paste0(prefix, txt)
  if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
  text(0.5, 0.5, txt, cex = cex.cor * r)
}

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#2. derive variables

@@

### Regression

```{echo=TRUE}

@@

library(stargazer)
stargazer(reg24, type = "text", digits = 2)

```{}

# residual plot
fit_res <-  data.frame(sm, fitted.value = fitted(reg24), residual = resid(reg24))

@@

```{r pressure, echo=FALSE}
plot(pressure)
```

## Conclusion

## References
European Commission (2010). Europe 2020 – A Strategy for smart, sustainable and inclusive growth. [online] Available at: http://eur-lex.europa.eu/LexUriServ/LexUriServ.do?uri=COM:2010:2020:FIN:EN:PDF [Accessed: 11.10.2016]

>>>>

# Students' Alcohol Consumption Impacts on Academic Results on the Example of a Secondary Scool Maths Class in Portugal

@@

2. Dataset Description and Descriptive Statistics

+ a. General Information
+ b. Family
+ c. Education 
+ d. Social Life and Health
+ e. Grades

3. Hypothesis Testing

4. Correlation Analysis and Regression

5. Interesting Findings

6. Conclusion

7. References

8. Appendices

## 1. Introduction 
Recent events in the world economy have lead to the decline in economic performance in many countries, amongst which is the decline of econommic growth in the European Union (EU).  As a response to these changes,  the EU has made it a priority to reverse this trend through investments that would provide new jobs and create a better educated workforce. Hence, the European Commission (EC) has set up a ‘Strategy for smart, sustainable and inclusive growth’ (European Comission, 2010), where a crucial part of investment is to contribute to the education of the next generations: The aim is to reduce the share of early school leavers to below 10% and ensure that at least 40% of the younger generation obtain a tertiary degree.

@@

## 2. Dataset Description and Descriptive Statistics 
```{r pressure, include=FALSE}
# Read csv file
sm <- read.csv("student-mat.csv", sep=';')

# Descriptive statistics
library(ggplot2)
library(gridExtra)

@@

df_gender <- data.frame(Gender=c("Female", "Male"), value=c(208, 187))
pie_gender <- ggplot(data=df_gender, aes(x="", y=value, fill=Gender)) + 
  geom_bar(width = 1, stat = "identity") + 
  coord_polar("y", start=0) + 
  scale_fill_brewer() + blank_theme + 
  theme(axis.text.x=element_blank()) +
  geom_text(aes(y = value/3 + c(0, cumsum(value)[-length(value)]), label = (value)), size=4)

@@

df_studyTime <- data.frame(Hours=c('<2 hours', "2 to 5 hours", "5 to 10 hours", ">10 hours"), value=c(105, 198, 65, 27))
#weekly study time (numeric: 1 - <2 hours, 2 - 2 to 5 hours, 3 - 5 to 10 hours, or 4 - >10 hours)

bar_studytime <- ggplot(data=df_studyTime, aes(x="", y=value, fill=Hours)) + 
  geom_bar(width = 1, stat = "identity") + 
  scale_fill_brewer() + blank_theme + 
  theme(axis.text.x=element_blank()) +
  geom_text(aes(y = value/3 + c(0, cumsum(value)[-length(value)]), label = (value)), size=4) +
  ggtitle('Study Time')

@@

df_school <- data.frame(Schools=c("Gabriel Pereira", "Mousinho da Silveira"), value=c(349,46))
bar_school <- ggplot(data=df_school, aes(x="", y=value, fill=Schools)) + 
  geom_bar(width = 1, stat = "identity") + 
  scale_fill_brewer() + blank_theme +
  theme(axis.text.x=element_blank()) +
  geom_text(aes(y = value/3 + c(0, cumsum(value)[-length(value)]), label = (value)), size=4)

@@

df_fjob <- data.frame(Job = c("Health", "Work at Home", "Teacher", "Servies", "Other"), value = c(18, 20, 29, 111, 217))
pie_fjob <- ggplot(data = df_fjob, aes(x="", y=value, fill=Job)) + 
  geom_bar(width = 1, stat = "identity") + 
  coord_polar("y", start=0) + 
  scale_fill_brewer() + blank_theme +
  theme(axis.text.x=element_blank()) +
  geom_text(aes(y = value/3 + c(0, cumsum(value)[-length(value)]), label = (value)), size=4) +
  ggtitle("Father's Job")

@@

df_mjob <- data.frame(Job = c("Health", "Teacher", "Work at Home", "Services", "Other"), value = c(34, 58, 59, 103, 141))
pie_mjob <- ggplot(data = df_mjob, aes(x="", y=value, fill=Job)) + 
  geom_bar(width = 1, stat = "identity") + 
  coord_polar("y", start=0) + 
  scale_fill_brewer() + blank_theme +
  theme(axis.text.x=element_blank()) +
  geom_text(aes(y = value/3 + c(0, cumsum(value)[-length(value)]), label = (value)), size=4) +
  ggtitle("Mother's Job")

@@

df_guardian <- data.frame(Type = c("Father", "Mother", "Other"), value = c(90, 273, 32))
pie_guardian <- ggplot(data = df_guardian, aes(x="", y=value, fill=Type)) + 
  geom_bar(width = 1, stat = "identity") + 
  coord_polar("y", start=0) + 
  scale_fill_brewer() + blank_theme +
  theme(axis.text.x=element_blank()) +
  geom_text(aes(y = value/3 + c(0, cumsum(value)[-length(value)]), label = (value)), size=4) +
  ggtitle("Guardian")
```

@@

#### a. General Information
General information about the students include gender, address and age. From the descriptive statistics (section 3) we can derive that the age from the students in this dataset ranges from 15 to 22 years with an average of 16.7 years. Out of the total of 395, there are 52.7% female and 47.3% male students. Most students would live in an urban environment, while approx. 20% live in an rural environment. 
```{r, echo=FALSE}
grid.arrange(pie_gender,pie_address, ncol=2)
```

#### b. Family  
Information in the students families includes the family size, educational background of their parents and their jobs, and the student's guardian. 
```{r, echo=FALSE}
grid.arrange(pie_fjob,pie_mjob,pie_guardian, ncol=3)
```

#### c. Education
The education section of this dataset covers the split between the two schools observed, why each student has chosen that particular school, and how long it takes each student to get to school. In addition, it provides information on the student's study time, extra educational support or classes taken and if the student is involved in extra curricular activities. Finally, there is data on their previous education and aspirations on whether to go into higher education. 
```{r, echo=FALSE}
grid.arrange(bar_school,bar_studytime, ncol=2)
```

#### d. Social Life and Health 
The social life section provieds information on activities the student is involved in outside of school. This includes how much leisure time they have, whether they are in a relationship and how often they go out. Moreover, there is data on how much alcohol students consume and in what health condition they are in.

#### e. Grades
The grades are split into three grades: the first period (G1), second period (G2) and final grade (G3). For the purpose of thy analysis, G1 and G2 will be neglected. In addition, there is information on whether they have failed to pass any classes. 
```{r, echo=FALSE}
pie_failures
```

## 3. Hypothesis Testing

## 4. Correlation Analysis and Regression
In this report, we would like to find factors that are related to students academic performance. We will perform a multivariate regression analysis and use step-wise variable selection to find significant variables related to our response variable (academic performance - "Score"). Prior to the regression analysis, we want to get a general idea about the relationship between the variables. There are several things we want to inspect:

a. Distribution of response variable

if there is skewness in the distribution, we should consider transforming the variable (i.e logarithmic)

b. Correlation between response variable and each explanatory variables

c. Correlation among explanatory variables

if there is a high correlation among explanatory variables, then there is possible multicollinearity problem

@@

### a. Distribution of Response Variable
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

sm[,"Score"] <- (sm[,"G1"] + sm[,"G2"] + sm[,"G3"])/3

ggplot(sm) + geom_density(aes(x=Score)) + labs(x = "Academic Score", y="% Frequency", title = "Academic Score of Secondary School Student")
```
From the graph, we can see that the distribution is slightly left-skewed. Generally, we would transform the variable to become more normally distributed. However, for the purpose of this project, we will not do any transfomation on the response variable and assume that the distribution is normal.

### b. Correlation between response variable and each explanatory variables
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...)
{
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  r <- abs(cor(x, y))
  txt <- format(c(r, 0.123456789), digits = digits)[1]
  txt <- paste0(prefix, txt)
  if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
  text(0.5, 0.5, txt, cex = cex.cor * r)
}

@@

### c. Correlation between response variable and each explanatory variables
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...)
{
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  r <- abs(cor(x, y))
  txt <- format(c(r, 0.123456789), digits = digits)[1]
  txt <- paste0(prefix, txt)
  if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
  text(0.5, 0.5, txt, cex = cex.cor * r)
}

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
#2. derive variables

@@

### Regression
```{echo=TRUE}

@@

library(stargazer)
stargazer(reg24, type = "text", digits = 2)
```

```{}
# residual plot
fit_res <-  data.frame(sm, fitted.value = fitted(reg24), residual = resid(reg24))

@@

```{r}
sm <- student.mat
for (i in 16:24) {                                  
  sm[,i] <- as.character(sm[,i])
  for (j in 1:nrow(sm)) {
    if (sm[j,i] == "yes") sm[j,i] <- 1
    if (sm[j,i] == "no") sm[j,i] <- 0
  }
  sm[,i] <- as.numeric(sm[,i])
}

# General test 
```{r}
t.test(sm[,"Walc"], sm[,"Dalc"])
library(ggplot2)
ggplot(data = sm) + geom_bar(aes(x=age))
ggplot(data = sm) + geom_bar(aes(x=sex))
ggplot(data = sm) + geom_bar(aes(x=Dalc))
ggplot(data = sm) + geom_bar(aes(x=Walc))
i <- ggplot(sm, aes(x = sm[,"age"], y = sm[,"sex"]))
i + geom_bin2d(binwidth = c(1,1)) #LABEL #Distribution of gender and age
i <- ggplot(sm, aes(x = sm[,"Dalc"], y = sm[,"Walc"]))
i + geom_bin2d(binwidth = c(1,1)) #LABEL #Distribution of Walc and Dalc
```

## Test and plot on gender and age
### age
```{r}
f <- ggplot(data =sm, aes(x = sm[,"Dalc"], y = sm[,"age"]))
f + geom_smooth(method = lm) #LABEL
f <- ggplot(data =sm, aes(x = sm[,"Walc"], y = sm[,"age"]))
f + geom_density(method = lm)  #LABEL
f + geom_smooth(na.rm = TRUE)
i <- ggplot(sm, aes(x = sm[,"age"], y = sm[,"sex"]))
i + geom_bin2d(binwidth = c(1,1))
i <- ggplot(sm, aes(x = sm[,"absences"], y = sm[,"Walc"]))
i + geom_bin2d(binwidth = c(1,1))
cor.test(sm[,"age"],sm[,"Dalc"])
cor.test(sm[,"age"],sm[,"Walc"])
```

### gender
```{r}
for (i in 2) {                                  
  sm[,i] <- as.character(sm[,i])
  for (j in 1:nrow(sm)) {
    if (sm[j,i] == "M") sm[j,i] <- 1
    if (sm[j,i] == "F") sm[j,i] <- 0
  }
  sm[,i] <- as.numeric(sm[,i])
}
cor.test(sm[,"sex"],sm[,"Walc"])
cor.test(sm[,"sex"],sm[,"Dalc"])
```

## Correlations tests 
As inputs have to be numeric it requires less lines to split the code that way instead of doing block per parameters analysed (Failures, paid, Medu, ... )
We could have done for loops here to calculate several correlations but it the results wasn't user friendly and doint it that way gave some useless correlations.
```{r}
sm <- as.numeric(sm)
sm1 <- subset(sm, sex == 1) #Create subsets with boys or girls
sm1 <- subset(sm, sex == 0)
```

### Romantic
```{r}
cor.test(sm[,"romantic"], sm[,"Dalc"])
cor.test(sm[,"romantic"], sm[,"Walc"])
```

### absenses
```{r}
cor.test(sm[,"absences"], sm[,"Dalc"])
cor.test(sm[,"absences"], sm[,"Walc"])
```

### Failures
```{r}
cor.test(sm[,"failures"], sm[,"Dalc"])
cor.test(sm[,"failures"], sm[,"Walc"])
```

### Boypart
```{r}
cor.test(sm1[,"failures"], sm1[,"Dalc"])
cor.test(sm1[,"failures"], sm1[,"Walc"])
```

### Medu
```{r}
cor.test(sm[,"Medu"], sm[,"Dalc"])
cor.test(sm[,"Medu"], sm[,"Walc"])
```

### Fedu
```{r}
cor.test(sm[,"Fedu"], sm[,"Dalc"])
cor.test(sm[,"Fedu"], sm[,"Walc"])
```

### Travel Time
```{r}
cor.test(sm[,"traveltime"], sm[,"Dalc"])
cor.test(sm[,"traveltime"], sm[,"Walc"])
```

### famrel
```{r}
cor.test(sm[,"famrel"], sm[,"Dalc"])
cor.test(sm[,"famrel"], sm[,"Walc"])
```

### goout
```{r}
cor.test(sm[,"goout"], sm[,"Dalc"])
cor.test(sm[,"goout"], sm[,"Walc"])
```

### internet
```{r}
cor.test(sm[,"internet"], sm[,"Dalc"])
cor.test(sm[,"internet"], sm[,"Walc"])
```

### famrel
```{r}
cor.test(sm[,"famrel"], sm[,"Dalc"])
cor.test(sm[,"famrel"], sm[,"Walc"])
```

### Grades
```{r}
cor.test(sm[,"G1"], sm[,"Dalc"])
cor.test(sm[,"G1"], sm[,"Walc"])
cor.test(sm[,"G2"], sm[,"Dalc"])
cor.test(sm[,"G2"], sm[,"Walc"])
cor.test(sm[,"G3"], sm[,"Dalc"])
cor.test(sm[,"G3"], sm[,"Walc"])
cor.test(sm[,"G3"], sm[,"goout"])
cor.test(sm[,"G3"], sm[,"failures"])
cor.test(sm[,"G3"], sm[,"absences"])
```

## Early Signs of Academic Failures 
### abscences
```{r}
cor.test(sm[,"absences"], sm[,"Dalc"])
cor.test(sm[,"absences"], sm[,"Walc"])
cor.test(sm1[,"absences"], sm1[,"Dalc"])
cor.test(sm1[,"absences"], sm1[,"Walc"])
cor.test(sm2[,"absences"], sm2[,"Dalc"])
cor.test(sm2[,"absences"], sm2[,"Walc"])
```

### Failures
```{r}
cor.test(sm[,"failures"], sm[,"Dalc"])
cor.test(sm[,"failures"], sm[,"Walc"])
cor.test(sm1[,"failures"], sm1[,"Dalc"])
cor.test(sm1[,"failures"], sm1[,"Walc"])
cor.test(sm2[,"failures"], sm2[,"Dalc"])
cor.test(sm2[,"failures"], sm2[,"Walc"])
```
